Research paper

### 2012-11-13 Research paper.tex
RP: Swapped background and project description
Added CLIN-abstract, RP: added sentence about phrase-based mt + references

When one of Starfleet's vessels meets an alien ship in areas where no man has gone before, they can (almost) always communicate with these aliens (almost) instantly; they speak to them in English, and these aliens answer in English. This is possible because, according to the Star Trek lore, the protagonists make use of a so-called Universal Translator. This Universal Translator can (1) recognize speech in every language, (2) translate what it has 'heard' to any other language, (3) make its results audible with speech synthesis and (4) learn itself new languages when needed. Unfortunately, like many technologies in Star Trek, this is far from reality: there is lots of active research at the first three tasks in particular, but no software currently in existence can do this as perfect as the Universal Translator. \\\indent
In this paper, I will give a detailed overview of the research project Colibri, which basically is an attempt to improve the technology needed for the second task: automatically translating text from one language to another. More specifically, the project investigates the identification and extraction of constructions in natural language, and how these can be used in Machine Translation. These constructions can be of various lengths, and, importantly, can also contain one or more gaps. The constructions are not identified on the basis of explicit 'human' knowledge about language and grammar ('linguistic theory'), but are distilled from large amounts of text ('text corpora') with the help of context-sensitive machine-learning techniques. The Colibri project is carried out by Maarten van Gompel and supervised by Antal van den Bosch.

%Beschrijving van ik ga doen.

@@

In this section I will try to give a short overview of the various approaches in machine translation, with special attention to the example-based approach, as that is the approach used in this project. This section is mainly based on %Maartens thesis.

@@

%Moet ik hier iets zeggen over phrase based?

@@

\bibitem[Van den Bosch \& Berk, 2009]{vdbb09}
Van den Bosch \& Berk, 2009

@@

>>>>

When one of Starfleet's vessels meets an alien ship in areas where no man has gone before, they can (almost) always communicate with these aliens (almost) instantly; they speak to them in English, and these aliens answer in English. This is possible because, according to the Star Trek lore, the protagonists make use of a so-called Universal Translator. This Universal Translator can (1) recognize speech in every language, (2) translate what it has 'heard' to any other language, (3) make its results audible with speech synthesis and (4) learn itself new languages when needed. Unfortunately, like many technologies in Star Trek, this is far from reality: there is lots of active research at the first three tasks in particular, but no software currently in existence can do this as perfect as the Universal Translator. \\\indent
In this paper, I will give a detailed overview of the research project Colibri, which basically is an attempt to improve the technology needed for the second task: automatically translating text from one language to another. More specifically, the project investigates the identification and extraction of constructions in natural language, and how these can be used in Machine Translation. These constructions can be of various lengths, and, importantly, can also contain one or more gaps. The constructions are not identified on the basis of explicit 'human' knowledge about language and grammar ('linguistic theory'), but are distilled from large amounts of text ('text corpora') with the help of context-sensitive machine-learning techniques. The Colibri project is carried out by Maarten van Gompel and supervised by Antal van den Bosch.
\\\indent
In section 2, I'll give a short overview of what has been achieved in machine translation, and in section 3 the current project will be described. This order was chosen to show that Colibri builds on previous research directly.

%Rest van de beschrijving.

@@

In this section I will try to give a short overview of the various approaches in machine translation, with special attention to the example-based approach, as that is the approach used in this project. This section is mainly based on \citet{vangompel09}

@@

On a final note, it should be added that although I focussed word-based translation (as described in \citet{stroppa07} and \citet{vdbb09}) in the previous examples,  \citet{vangompel09}, \citet{vangompelea09} and \citet{vangompel11} showed that even better results can be achieved using phrases of variable length. Investigating the advantage of multi-word translation units is one of the main goals of the Colibri project, as will be explained in the next section.

@@

\bibitem[Van den Bosch \& Berck, 2009]{vdbb09}
Van den Bosch, A., \& Berck, P. (2009). Memory-based machine translation and language
modeling. The Prague Bulletin of Mathematical Linguistics, 91, 17\-26.

@@

\bibitem[Van Gompel, 2009]{vangompel09}
Van Gompel, M. Phrase-based memory-based machine translation. Unpublished
master's thesis, Tilburg University.

\bibitem[Van Gompel et al., 2009]{vangompelea09}
Van Gompel, M., Van den Bosch, A., \& Berck, P. (2009). Extending memory-based
machine translation to phrases. In M. Forcada \& A. Way (Eds.), Proceedings of the
third workshop on example-based machine translation (pp. 79\-86). Dublin, Ireland.

\bibitem[Van Gompel et al., 2011]{vangompel11}
Van Gompel, M., Van den Bosch, A., \& Berck, P. (2011). Extending memory-based
machine translation to phrases. In T. Markus, P. Monachesi, \& E. Westerhout (Eds.),
Computational linguistics in the netherlands 2010: Selected papers from the twentieth
clin meeting (pp. 45\-58). Utrecht, the Netherlands: LOT.

\bibitem[Stroppa et al., 2007]{stroppa07}
Stroppa, N., Van den Bosch, A., \& Way, A. (2007). Exploiting source similarity for SMT
using context-informed features. In A. Way \& B. Gawronska (Eds.), Proceedings of the
11th international conference on theoretical issues in machine translation (tmi 2007)
(pp. 231\-240). Skövde, Sweden.
